Combinatorics of the geometry of Wilson loop diagrams

### 2019-01-08 General WLD combinatorics.tex
Updates to cyclic intervals lemma

Since Corollary \ref{uniqueproppartitioncor} allows us to decompose any admissible Wilson loop diagram into a collection of maximal exact subdiagrams, in this section we examine the matroid properties of exact subdiagrams more closely. In particular, we show that given an exact subdiagram $(P, V(P))$, it can be written as the contraction of the matroid $M(W)$ by the complementary propagator flat, $F(P^c)$. Furthermore, if $(P, V(P))$ is a maximal exact subdiagrams, its complementary propagator flat is a cyclic flat. Finally, we show that matroids associated to exact subdiagrams are uniform (Theorem \ref{exactuniformthm}).

We begin by proving some useful facts about flats of matroids associated to admissible Wilson loop diagrams.

@@

\begin{proof}
We will prove the result by induction on the number of propagators.  If $W$ has one propagator then the result is immediate.  Now suppose $W$ has more than one propagator.  Since non-supporting vertices play no role in Algorithm~\ref{alg:put GN on WLD}, it suffices to prove the result for $W$ with $F(\emptyset) = \emptyset$.  Then by Lemma~\ref{lem sian}, $W$ has at least one of the four situations illustrated in Figure~\ref{fig 3 cases}.

@@

\textbf{Left two cases}: These cases behave essentially identically (except when $j$ or $j+1$ are not in the support of $p$, which can occur in the second case only; see below) so we handle the majority of the proof for these two cases simultaneously. Let $1\leq a\leq 3$ be the number of unsupported vertices inside $q$ in $V$; so these vertices are $l, \ldots, l+a-1$.  Write $p=(m,m+2)$ where $m\in \{i, i+1, l\}$.  Note that $l, \ldots, l+a-1$ are all in the support of $p$.

@@

\textbf{Third case}: In this case there are no unsupported vertices $l, l+1, \ldots$ inside $q$.  Again write $p=(m, m+2)$ where $m\in \{j, j+1, j+2\}$. We proceed as in the previous cases, by computing $I^{(W)}_w$ for vertices $w$ in roughly increasing order of difficulty.

@@

\begin{align*}
  I_{j+1}^{(W)}(r) &= \begin{cases}
    j+1 & \text{if } r=q\\
    j+2 & \text{if } r=p\\
    I_{j+3}^{(V)}(r) & \text{if } r\neq p,q
  \end{cases}\\
  I_m^{(W)}(r) &= \begin{cases}
    m & \text{if $r=p$ and $q$ not supported on $m$}\\
    m+1 & \text{if $r=p$ and $q$ supported on $m$}\\
    I_{m}^{(V)}(r) & \text{if } r\neq p
  \end{cases}\\
  I_{m+1}^{(W)}(r) & = \begin{cases}
    m+1 & \text{if $r=p$ and $q$ not supported on $m+1$} \\
    m+2 & \text{if $r=p$ and $q$ supported on $m+1$} \\
    I_{m+1}^{(V)}(r) & \text{if } r\neq p,q
  \end{cases}\\
  I_{m+2}^{(W)}(r) & = \begin{cases}
    m+2 & \text{if } r=p \\
    I_{m+3}^{(V)}(r) & \text{if } r\neq p
  \end{cases}\\
  I_{m+3}^{(W)}(r) & = \begin{cases}
    m+3 & \text{if } r=p \\
    I_{m+4}^{(V)}(r) & \text{if } r\neq p
  \end{cases}
\end{align*}
Note that $I_{m+2}^{(V)}(r) = I_{m+3}^{(V)}(r)$ for all propagators $r$ in $V$, and that if $j+1\not\in\{m, m+1, m+2, m+3\}$ then $j+2$ and $j+3$ are unsupported in $V$ so in that case $I_{j+2}^{(V)}(r) = I_{j+3}^{(V)}(r) = I_{j+4}^{(V)}(r)$ for $r$ in $V$.

@@

Let $W'$ be the diagram obtained from $W$ by removing only the propagator $p$.  Then we have
\begin{align*}
  I_i^{(W)} \backslash \{i\} & = I_{i+1}^{(W')} \\
  I_{i+1}^{(W)} \backslash \{i+1\} & = I_{i+2}^{(W')}
\end{align*}
since in both cases the algorithm in $W'$ proceeds identically to that in $W$ after assigning $p$. Since $k \neq i+1$, we have $k\in I_{i+1}^{(W')}\setminus\{i+1\}$ but $k\not\in I_{i+2}^{(W')}$, contradicting the minimiality of $(W, i, k)$

>>>>

Since Corollary \ref{uniqueproppartitioncor} allows us to decompose any admissible Wilson loop diagram into a collection of maximal exact subdiagrams, in this section we examine the matroid properties of exact subdiagrams more closely. In this section, we show prove two results:
\begin{enumerate}
\item The matriod assoicated to an exact subdiagram of $W$ can be written as a contraction of the matroid $M(W)$ by the complementary propagator flat (Theorem \ref{exact diagrams contractions}). 
\item The matroid associated to an exact subdiagram is uniform (Theorem \ref{exactuniformthm} ).
\end{enumerate}

We begin by proving some useful facts about propagator flats, and flats of matroids associated to admissible Wilson loop diagrams more generally.

@@

\begin{proof}
We will prove the result by induction on the number of propagators.  If $W$ has one propagator then the result is immediate.  Now suppose $W$ has more than one propagator.  Since non-supporting vertices have no effect on the Grassmann necklace algorithm, it suffices to prove the result for $W$ with $F(\emptyset) = \emptyset$.  Then by Lemma~\ref{lem sian}, $W$ has at least one of the four situations illustrated in Figure~\ref{fig 3 cases}.

@@

\textbf{Left two cases}: These cases behave essentially identically (except when $j$ or $j+1$ are not in the support of $p$, which can occur in the second case only; see below) so we handle the majority of the proof for these two cases simultaneously. Let $1\leq a\leq 3$ be the number of non-supporting vertices inside $q$ in $V$; so these vertices are $l, \ldots, l+a-1$.  Write $p=(m,m+2)$ where $m\in \{i, i+1, l\}$.  Note that $l, \ldots, l+a-1$ are all in the support of $p$.

@@

\textbf{Third case}: In this case there are no non-supporting vertices $l, l+1, \ldots$ inside $q$.  Again write $p=(m, m+2)$ where $m\in \{j, j+1, j+2\}$. We proceed as in the previous cases, by computing $I^{(W)}_w$ for vertices $w$ in roughly increasing order of difficulty.

@@

\begin{align*}
  I_{j+1}^{(W)}(r) &= \begin{cases}
    j+1 & \text{if } r=q\\
    j+2 & \text{if } r=p\\
    I_{j+3}^{(V)}(r) & \text{if } r\neq p,q
  \end{cases}\\
  I_m^{(W)}(r) &= \begin{cases}
    m & \text{if $r=p$ and $q$ not supported on $m$}\\
    m+1 & \text{if $r=p$ and $q$ supported on $m$}\\
    I_{m}^{(V)}(r) & \text{if } r\neq p
  \end{cases}\\
  I_{m+1}^{(W)}(r) & = \begin{cases}
    m+1 & \text{if $r=p$ and $q$ not supported on $m+1$} \\
    m+2 & \text{if $r=p$ and $q$ supported on $m+1$} \\
    I_{m+1}^{(V)}(r) & \text{if } r\neq p,q
  \end{cases}\\
  I_{m+2}^{(W)}(r) & = \begin{cases}
    m+2 & \text{if } r=p \\
    I_{m+3}^{(V)}(r) & \text{if } r\neq p
  \end{cases}\\
  I_{m+3}^{(W)}(r) & = \begin{cases}
    m+3 & \text{if } r=p \\
    I_{m+4}^{(V)}(r) & \text{if } r\neq p
  \end{cases}
\end{align*}
Note that $I_{m+2}^{(V)}(r) = I_{m+3}^{(V)}(r)$ for all propagators $r$ in $V$, and that if $j+1\not\in\{m, m+1, m+2, m+3\}$ then $j+2$ and $j+3$ are non-supporting vertices in $V$, so in that case $I_{j+2}^{(V)}(r) = I_{j+3}^{(V)}(r) = I_{j+4}^{(V)}(r)$ for $r$ in $V$.

@@

Let $W'$ be the diagram obtained from $W$ by removing only the propagator $p$.  Then we have
\begin{align*}
  I_i^{(W)} \backslash \{i\} & = I_{i+1}^{(W')} \\
  I_{i+1}^{(W)} \backslash \{i+1\} & = I_{i+2}^{(W')}
\end{align*}
since in both cases the algorithm in $W'$ proceeds identically to that in $W$ after assigning $p$. Since $k \neq i+1$, we have $k\in I_{i+1}^{(W')}\setminus\{i+1\}$ but $k\not\in I_{i+2}^{(W')}$, contradicting the minimiality of $(W, i, k)$.
